Comparison Word2Vec Analysis

### 2018-08-28 Hills2009Replication.Rmd
Presentation prep

# Methods

```{r echo=FALSE, message=FALSE, warning=FALSE} 
library(purrr) # fp library
library(readr) # read text data files (csv, etc.)
library(ggplot2) # plotting
library(langcog) # useful things used by Language and Cognition Lab
library(dplyr) # for working with datafram-like objects
library(tidyr) # for data tidying
library(wordbankr) # access to wordbank data
library(stringr) # common string operations
library(broom) # converts R function outputs to tidy data frames
library(igraph) # network analysis and visualization
library(knitr) # general-purpose tool for dynamic report generation
library(kableExtra) # table formatting
library(lemon) # functions for working w/ ggplot2 and knitr extensions
library(tibble) # modern dataframes
library(tidygraph)
library(doParallel)
library(networkD3)
library(ForceAtlas2)

@@

full_vocab <- make_vocab_dataframe(lang="English (American)",
                              lang_form = "WS",
                              lex_class = "nouns") %>%
  trim_all_unilemma() %>%
  trim_all_definition()
```

@@

```{r warning=FALSE}
features <- load_features()

vocab <- full_vocab %>%
  filter(age <= 30) %>%
  filter(uni_lemma %in% features$uni_lemma) %>%
  distinct(uni_lemma, .keep_all = TRUE)
```

Only the words for which feature data are available are considered, an overlap of `r nrow(vocab)` in our data compared to Hills' 130.

@@

full_network <- feature_network(vocab, features, shared_threshold = 1,
                                feature_types = c(perceptual_features,
                                                  conceptual_features)) %>%
                  set.graph.attribute("type", "both") %>%
                  set.graph.attribute("age", 30)

perceptual_network <- feature_network(vocab, features, shared_threshold = 1,
                                      feature_types = perceptual_features) %>%
                        set.graph.attribute("type", "perceptual") %>%
                        set.graph.attribute("age", 30)

conceptual_network <- feature_network(vocab, features, shared_threshold = 1,
                                      feature_types = conceptual_features) %>%
                        set.graph.attribute("type", "conceptual") %>%
                        set.graph.attribute("age", 30)
```

@@

## The Full Network

### Network statistics and small-world analyses

```{r}
# testing dense layout
source(paste(getwd(),"/helpers/all_helper.r",sep = ""), chdir = T)

plot_network(full_network %E>% filter(shared >= 2), remove_isolated = TRUE, labels=TRUE, frame=TRUE)
```

The full network is shown in the following figure, for w = 1, 2, 3, and 4.

```{r out.width = '100%'}
full_networks = list()
for (w in 1:4) {
  full_networks[[w]] = full_network %E>%
    filter(shared >= w)
}

#par(mfcol=c(2, 2), mar=c(0.5,0.5,0.5,0.5))
for (network in full_networks) {
  plot_network(network, remove_isolated = FALSE, labels=TRUE, frame=TRUE)
}
```

## Basic Vocab and Feature Properties

## Basic Network Properties

Some basic properties of networks built from perceptual and conceptual features.
```{r, results='asis'}
max_w <- 4

network_properties <- function(networks) {
  return( tibble(
  name = map_chr(networks, ~ .x$name),
  type = map_chr(networks, ~ .x$type),
  age = map_dbl(networks, ~ .x$age) %>%
    as.integer(),
  w = map_dbl(networks, ~ .x$w) %>%
    as.integer(),
  vertices = map_int(networks, vcount),
  edges = map_dbl(networks, ecount) %>%
    as.integer(), # for some reason igraph ecount returns double
  avg_degree = map(networks, degree) %>%
    map_dbl(mean),
  avg_shortest_path = map_dbl(networks, mean_distance),
  diameter = map_dbl(networks, diameter) %>%
    as.integer(),
  clustering_coefficient = map_dbl(networks, transitivity),
  density = edges / (vertices * (vertices - 1) / 2),
  avg_shortest_path_random = 0, # either estimate or calculate values for random nets
  diameter_random = 0,
  clustering_coefficient_random = density,
  isolates = map_dbl(networks, function (g) sum(degree(g) == 0)),
  components = map_dbl(networks, count_components) %>%
    as.integer() - isolates))
}

full_networks <- list()
perceptual_networks <- list()
conceptual_networks <- list()
for (w in 1:max_w) {
  full_networks[[w]] <- full_network %>%
    activate(what = edges) %>%
    filter(shared >= w) %>%
    set.graph.attribute("w", w)

  perceptual_networks[[w]] <- perceptual_network %>%
    activate(what = edges) %>%
    filter(shared >= w) %>%
    set.graph.attribute("w", w)

  conceptual_networks[[w]] <- conceptual_network %>%
    activate(what = edges) %>%
    filter(shared >= w) %>%
    set.graph.attribute("w", w)
}

networks <- full_networks %>%
  append(perceptual_networks) %>%
  append(conceptual_networks)

our_network_properties <- network_properties(networks)

our_network_properties %>%
  select(type, w, vertices, edges, avg_degree, diameter, avg_shortest_path, clustering_coefficient) %>%
  kable(format="html",
        escape = FALSE,
        col.names = c("Type", "w", "V", "E", "Average degree", "Diameter", "Average length", "Clustering coefficient"),
        digits = 2,
        align = rep("l", 8)) %>%
  kable_styling(full_width = F) %>%
  collapse_rows(columns = 1:3, valign = "top")
```

We can compare these networks to those from Hills, noting that we have a different set of words/nodes but that the induced subgraph of those words shared between our analysis and that of Hills should be the same.
```{r}
hills_network_properties <- tibble( # table from Hills 2009
  type = c(rep("all", times=4), rep("perceptual", times=4), rep("conceptual", 4)),
  w = c(1:max_w, 1:max_w, 1:max_w),
  clustering_coefficient = c(0.55, 0.54, 0.58, 0.60, 0.54, 0.54, 0.62, 0.62, 0.88, 0.59, 0.38, 1.00),
  avg_shortest_path = c(1.72, 2.57, 4.96, 3.22, 1.76, 2.69, 5.02, 2.55, 3.08, 1.59, 5.02, 1.00),
  density = c(0.29, 0.10, 0.04, 0.02, 0.27, 0.08, 0.03, 0.01, 0.05, 0.01, 0.0, 0.0),
  components = c(1, 1, 4, 10, 1, 2, 8, 11, 7, 14, 7, 2),
  isolates = c(0, 6, 20, 47, 0, 10, 31, 69, 33, 81, 113, 126)
)

# convert contents to text
display_our_properties <- our_network_properties %>% mutate_all(funs(format(., digits=2, trim=TRUE)))
display_hills_properties <- hills_network_properties %>% mutate_all(funs(paste("(", ., ")", sep='')))

# combine text tables for display
display_properties <- bind_rows(display_our_properties %>% mutate(row = 1:n()),
                                display_hills_properties %>% mutate(row = 1:n())) %>%
  group_by(row) %>%
  summarise_if(is.character, funs(paste(., collapse=" "))) %>%
  arrange(row)

# display combined properties
select(hills_network_properties, type, w) %>%
  bind_cols(select(display_properties, clustering_coefficient, avg_shortest_path,
                   density, components, isolates)) %>%
  kable(format="html",
        escape = FALSE,
        col.names = c("Type", "w", "Clustering Coefficient", "Average length", "Density", "Components", "Isolates"),
        align = rep("l", 7)) %>%
  kable_styling(full_width = F) %>%
  collapse_rows(columns = 1:2, valign = "top")
```

# Clusters

## Visualize Networks
```{r out.width = '100%'}
networks <- perceptual_networks[2]
```

## Clique Percolation

Initial Clique Analysis
```{r out.width = '100%'}
trimmed_networks <- map(networks, function (g) delete.vertices(g, degree(g)==0))

# cl <- makeCluster(3)
# registerDoParallel(cl)
# getDoParWorkers()
## Plot with the new colors
for (net in networks) {
  for (k in 3:4) {
    clusters <- clique.community.opt(net,k)
    plot_network(net, clusters = clusters, labels=FALSE)
  }
}
# stopCluster(cl)
```

Dynamic Visualization
```{r}
conceptual_clusters <- membership(cluster_walktrap(conceptual_network))

# Convert to object suitable for networkD3
conceptual_d3 <- igraph_to_networkD3(conceptual_network, group = conceptual_clusters)

# Create force directed network plot
forceNetwork(Links = conceptual_d3$links, Nodes = conceptual_d3$nodes,
             Source = 'source', Target = 'target', NodeID = 'name',
             Group = 'group')
```

>>>>

# Preparation

This analysis requires a working copy of CFinder and the CFinder license in the same directory (necessary for running CFinder). The following must point to the command line executable:

```{r}
cfinder_program <- "./CFinder_commandline_mac"
```

```{r echo=FALSE, message=FALSE, warning=FALSE} 
library(purrr) # fp library
library(readr) # read text data files (csv, etc.)
library(ggplot2) # plotting
library(langcog) # useful things used by Language and Cognition Lab
library(dplyr) # for working with datafram-like objects
library(tidyr) # for data tidying
library(wordbankr) # access to wordbank data
library(stringr) # common string operations
library(broom) # converts R function outputs to tidy data frames
library(igraph) # network analysis and visualization
library(knitr) # general-purpose tool for dynamic report generation
library(kableExtra) # table formatting
library(lemon) # functions for working w/ ggplot2 and knitr extensions
library(tibble) # modern dataframes
library(tidygraph)
library(doParallel)
library(networkD3)
library(formattable)
library(viridis)

@@

full_vocab <- make_vocab_dataframe(lang="English (American)",
                                   lang_form = "WS",
                                   lex_class = "nouns") %>%
  trim_all_unilemma() %>%
  trim_all_definition() %>%
  arrange(age) %>%            # keep first age for duplicates arising from trimming
  distinct(uni_lemma, .keep_all = TRUE) %>%
  select(uni_lemma, definition, age, category)
```

@@

```{r warning=FALSE, message=FALSE}
vocab <- full_vocab %>%
  filter(age <= 30) %>%
  filter_to_features()
```

Only the words aquired by 30 months for which feature data are available are considered, an overlap of `r nrow(vocab)` in our data compared to Hills' 130.

@@

feature_links <-
  make_feature_links(vocab = vocab, feature_types = union(perceptual_features, conceptual_features))
perceptual_links <-
  make_feature_links(vocab = vocab, feature_types = perceptual_features)
conceptual_links <-
  make_feature_links(vocab = vocab, feature_types = conceptual_features)

feature_network <- graph_from_data_frame(feature_links, vertices = vocab, directed = FALSE) %>%
  as_tbl_graph() %E>%
  mutate(weight = shared) %>%
  set.graph.attribute("type", "Features") %>%
  set.graph.attribute("age", 30)

perceptual_network <- graph_from_data_frame(perceptual_links, vertices = vocab, directed = FALSE) %>%
  as_tbl_graph() %E>%
  mutate(weight = shared) %>%
  set.graph.attribute("type", "Perceptual") %>%
  set.graph.attribute("age", 30)

conceptual_network <- graph_from_data_frame(conceptual_links, vertices = vocab, directed = FALSE) %>%
  as_tbl_graph() %E>%
  mutate(weight = shared) %>%
  set.graph.attribute("type", "Conceptual") %>%
  set.graph.attribute("age", 30)
```

@@

## Network statistics and small-world analyses

The network are shown in the following figure, for w = 1, 2, 3, and 4.

```{r out.width = '25%', fig.asp=1}
par(oma=c(0,0,0.2,0))

for (network in list(feature_network, perceptual_network, conceptual_network)) {
  for (w in 1:4) {
    net <- network %E>% filter(shared >= w)
    layout <- layout_with_fr(net)
    plot_vocab_network(net, plot_title = title(str_glue("{net$type} w={w}"), cex.main=2), frame=FALSE, layout = layout, labels="")
  }
}
```

Some basic properties of networks built from perceptual and conceptual features.

```{r, results='asis'}
network_properties <- function(network) {
  return( tibble(
    type = network$type,
    age = network$age,
    w = network$w,
    vertices = vcount(network),
    edges = ecount(network),
    avg_degree = mean(degree(network)),
    avg_distance = mean_distance(network),
    diameter = diameter(network),
    clustering_coef = transitivity(network),
    density = edges / (vertices * (vertices - 1) / 2),
    clustering_coefficient_random = density,
    isolates = sum(degree(network) == 0),
    components = count_components(network) - isolates
  ))
}

max_w <- 4

feature_networks <- map(1:max_w, function(w) {
  feature_network %>%
    filter(shared >= w) %>%
    set.graph.attribute("w",w)})
perceptual_networks <- map(1:max_w, function(w) {
  perceptual_network %>%
    filter(shared >= w) %>%
    set.graph.attribute("w",w)})
conceptual_networks <- map(1:max_w, function(w) {
  conceptual_network %>%
    filter(shared >= w) %>%
    set.graph.attribute("w",w)})

networks <- feature_networks %>%
  append(perceptual_networks) %>%
  append(conceptual_networks)

properties <- bind_rows(map(networks, network_properties))
```

We can compare these networks to those from Hills, noting that we have a different set of words/nodes but that the induced subgraph of those words shared between our analysis and that of Hills should be the same.

```{r}
properties %>%
  mutate(h_clustering_coef = c(0.55, 0.54, 0.58, 0.60, 0.54, 0.54, 0.62, 0.62, 0.88, 0.59, 0.38, 1.00),
         h_avg_distance = c(1.72, 2.57, 4.96, 3.22, 1.76, 2.69, 5.02, 2.55, 3.08, 1.59, 5.02, 1.00),
         h_density = c(0.29, 0.10, 0.04, 0.02, 0.27, 0.08, 0.03, 0.01, 0.05, 0.01, 0.0, 0.0),
         h_components = c(1, 1, 4, 10, 1, 2, 8, 11, 7, 14, 7, 2),
         h_isolates = c(0, 6, 20, 47, 0, 10, 31, 69, 33, 81, 113, 126)) %>%
  select(type, age, w,
         clustering_coef, h_clustering_coef,
         avg_distance, h_avg_distance,
         density, h_density,
         components, h_components,
         isolates, h_isolates) %>%
  kable(format="html",
        escape = FALSE,
        col.names = c("Type", "age", "w",
                      "Clustering Coefficient", "",
                      "Average length", "",
                      "Density", "",
                      "Components", "",
                      "Isolates", ""),
        align = c("l", "l", "l", rep(c("r", "l"), 5)),
        digits = 2) %>%
  kable_styling(full_width = F) %>%
  column_spec(seq(5,13,2), color = "grey") %>%
  collapse_rows(columns = 1:2, valign = "top")
```

## Clique Percolation Clustering

### Full Network

Here we find parameter values for clique percolation according to the "principled approach" given in Hills, raising k for a given w until the the second largest cluster is greater than half the size of the largest, and then choosing the k for which this gives the most clusters.

```{r}
best_count <-  0
for (w in 1:5) {
  result <- cluster_cfinder(feature_network %E>% filter(shared >= w),
                            cfinder_program = cfinder_program)
  clusters <- result$clusters
  k <- result$k
  if (length(clusters) > best_count) {
    best_w <- w
    best_k <- k
    best_count <- length(clusters)
    best_clusters <- clusters
  }
}
```

This gives values of w=`r best_w` and k=`r best_k`, as opposed to Hills' w=3 and k=3, and gives the following clusters:

```{r}
title <- str_glue("Feature Network Clique Percolation Clusters, w = {best_w} and k = {best_k}")
kable_clusters(best_clusters, title)
```

Hills' values of w=3 and k=3 also give reasonable clusters:

```{r}
clusters <- cluster_cfinder(feature_network %E>% filter(shared >= 3),
                            k = 3,
                            cfinder_program = cfinder_program)$clusters
kable_clusters(clusters, "Feature Network Clique Percolation Clusters, w = 3 and k = 3")
```

Note that this difference may be due to the minor differences between our networks. The largest cluster produced by w=3, k=3 is only one larger than twice the size of the second largest, and had it been smaller these parameters would have returned more clusters (and thus been selected) than w=`r best_w` and k=`r best_k`. However, these minor differences do lead to substantially different clusterings.

### Perceptual Network

Here we find parameter values for clique percolation on the perceptual network.

```{r}
best_count <-  0
for (w in 1:5) {
  result <- cluster_cfinder(perceptual_network %E>% filter(shared >= w),
                            cfinder_program = cfinder_program)
  clusters <- result$clusters
  k <- result$k
  if (length(clusters) > best_count) {
    best_w <- w
    best_k <- k
    best_count <- length(clusters)
    best_clusters <- clusters
  }
}
```

This gives values of w=`r best_w` and k=`r best_k`, as opposed to Hills' w=2 and k=5, and gives the following clusters:

```{r}
title <- str_glue("Perceptual Network Clique Percolation Clusters, w = {best_w} and k = {best_k}")
kable_clusters(best_clusters, title)
```

Hills' values of w=2 and k=5 also give reasonable clusters:

```{r}
clusters <- cluster_cfinder(perceptual_network %E>% filter(shared >= 2),
                            k = 5,
                            cfinder_program = cfinder_program)$clusters
kable_clusters(clusters, "Feature Network Clique Percolation Clusters, w = 2 and k = 5")
```

### Conceptual Network

Here we find parameter values for clique percolation on the conceptual network.

```{r}
best_count <-  0
for (w in 1:5) {
  result <- cluster_cfinder(conceptual_network %E>% filter(shared >= w),
                            cfinder_program = cfinder_program)
  clusters <- result$clusters
  k <- result$k
  if (length(clusters) > best_count) {
    best_w <- w
    best_k <- k
    best_count <- length(clusters)
    best_clusters <- clusters
  }
}
```

This gives values of w=`r best_w` and k=`r best_k`, matching Hills' w=1 and k=3, and gives the following clusters:

```{r}
title <- str_glue("Conceptual Network Clique Percolation Clusters, w = {best_w} and k = {best_k}")
kable_clusters(best_clusters, title)
```

### 2018-08-28 Main.Rmd
Presentation prep

---
title: "Word2Vec Analysis"
author: "Isaac Scheinfeld"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: default
---
# Networks

@@

layout <- layout_network(network %E>% mutate(weights = w2v_similarity))
```

@@

```{r out.width = '100%', fig.height = 5, fig.width = 5}
source(paste(getwd(),"/helpers/all_helper.r",sep = ""), chdir = T)

@@

for (s in seq(0.35, 0.5, 0.05)) {
  plot_network(network %E>% filter(w2v_similarity > s),
               layout = layout,
               labels=TRUE,
               title=str_glue("Threshold = {s}"))
}

@@

Test igraph clustering algorithms

```{r fig.height = 5, fig.width = 5}
source(paste(getwd(),"/helpers/all_helper.r",sep = ""), chdir = T)

# cluster_edge_betweenness - doesn't cluster reasonably, treats weights as distances?

# cluster_infomap - returns one cluster

# cluster_label_prop - returns one cluster

# cluster_optimal - too computationally intensive for laptop, may not be possible (NP)

plot_network(network, layout = layout,
             clusters = cluster_fast_greedy(network, weights = E(network)$w2v_similarity)$membership,
             labels=TRUE, frame=TRUE)

plot_network(network, layout = layout,
             clusters = cluster_leading_eigen(network, weights = E(network)$w2v_similarity)$membership,
             labels=TRUE, frame=TRUE)

plot_network(network, layout = layout,
             clusters = cluster_louvain(network, weights = E(network)$w2v_similarity)$membership,
             labels=TRUE, frame=TRUE)

plot_network(network, layout = layout,
             clusters = cluster_spinglass(network, weights = E(network)$w2v_similarity)$membership,
             labels=TRUE, frame=TRUE)

plot_network(network, layout = layout,
             clusters = cluster_walktrap(network, weights = E(network)$w2v_similarity)$membership,
             labels=TRUE, frame=TRUE)
```

>>>>

---
title: "Comparison Word2Vec Analysis"
author: "Isaac Scheinfeld"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: default
---
# Preparation

@@

layout <- layout_network(network %E>% mutate(weights = w2v_similarity))

network <- network %N>%
  mutate(x = layout[,1]) %>%
  mutate(y = layout[,2])
```

@@

```{r out.width = '100%', dpi=300}
source(paste(getwd(),"/helpers/all_helper.r",sep = ""), chdir = T)

@@

for (s in seq(0.35, 0.5, 0.05)) {
  plot_network(network %E>% filter(w2v_similarity > s),
               layout = layout,
               labels=TRUE,
               frame=FALSE,
               title=str_glue("Threshold = {s}"))
}

@@

## Clustering algorithms

```{r out.width='100%', fig.asp=1/2, dpi=300}
# cluster_edge_betweenness - doesn't cluster reasonably, treats weights as distances?
# 
# cluster_infomap - returns one cluster
# 
# cluster_label_prop - returns one cluster
# 
# cluster_optimal - too computationally intensive for laptop, may not be possible (NP)

source(paste(getwd(),"/helpers/all_helper.r",sep = ""), chdir = T)

par(mfrow=c(1, 2), mar=c(2,0.4,0.8,0.4)) # row x col, column first

feature_network <- network %E>%
  mutate_all(funs(replace(., is.na(.), 0))) %>%
  mutate(features = perceptual + conceptual) %>%
  filter(features > 0)

# cluster_fast_greedy
plot_network(network, layout = layout,
             clusters = cluster_fast_greedy(network,
                                            weights = E(network)$w2v_similarity)$membership,
             labels=TRUE, frame=TRUE)

plot_network(feature_network, layout = layout,
             clusters = cluster_fast_greedy(feature_network,
                                            weights = E(feature_network)$features)$membership,
             labels=TRUE, frame=TRUE)

# cluster_leading_eigen
plot_network(network, layout = layout,
             clusters = cluster_leading_eigen(network,
                                              weights = E(network)$w2v_similarity)$membership,
             labels=TRUE, frame=TRUE)

plot_network(feature_network, layout = layout,
             clusters = cluster_leading_eigen(feature_network,
                                              weights = E(feature_network)$features)$membership,
             labels=TRUE, frame=TRUE)

# cluster_louvain
plot_network(network, layout = layout,
             clusters = cluster_louvain(network,
                                        weights = E(network)$w2v_similarity)$membership,
             labels=TRUE, frame=TRUE)

plot_network(feature_network, layout = layout,
             clusters = cluster_louvain(feature_network,
                                        weights = E(feature_network)$features)$membership,
             labels=TRUE, frame=TRUE)

# cluster_spinglass
plot_network(network, layout = layout,
             clusters = cluster_spinglass(network,
                                          weights = E(network)$w2v_similarity)$membership,
             labels=TRUE, frame=TRUE)

plot_network(feature_network, layout = layout,
             clusters = cluster_spinglass(feature_network,
                                          weights = E(feature_network)$features)$membership,
             labels=TRUE, frame=TRUE)

# cluster_walktrap
plot_network(network, layout = layout,
             clusters = cluster_walktrap(network,
                                         weights = E(network)$w2v_similarity)$membership,
             labels=TRUE, frame=TRUE)

plot_network(feature_network, layout = layout,
             clusters = cluster_walktrap(feature_network,
                                         weights = E(feature_network)$features)$membership,
             labels=TRUE, frame=TRUE)
```

```{r out.width='100%', fig.asp=1/2, dpi=300}
# cluster_walktrap
for (age_limit in seq(16, 30, by=2)) {
  net_at_age <- network %N>% filter(age <= age_limit) %E>% mutate(weights = w2v_similarity)

  plot_network(net_at_age,
               clusters = cluster_walktrap(net_at_age,
                                           weights = E(net_at_age)$w2v_similarity)$membership,
               labels = TRUE,
               frame = TRUE)
}
```

```{r}
feature_net <- network %E>% mutate(weights = perceptual) %E>% filter(weights > 0) %N>% filter(!node_is_isolated())
feature_layout <- layout_network(feature_net)

feature_net <- feature_net %N>%
  mutate(x = feature_layout[,1]) %>%
  mutate(y = feature_layout[,2])

```{r out.width='100%', fig.asp=1/2, dpi=300}
# cluster_walktrap
for (age_limit in seq(16, 30, by=2)) {
  net_at_age <- feature_net %N>% filter(age <= age_limit) %E>% mutate(weights = perceptual)

  plot_network(net_at_age,
               clusters = cluster_walktrap(net_at_age,
                                           weights = E(net_at_age)$perceptual)$membership,
               labels = TRUE,
               frame = TRUE)
}
```
